---
title: "Dplyr Homework"
output: html_notebook
---

```{r}
library(CodeClanData)
drinks_content
library(dplyr)

```



##Q1 dimensions; columns = 18, rows = 242, types = chr, num, names = below;, Vitamin C % etc, 1 NA value found

```{r}
glimpse(drinks_content)
names(drinks_content)
sum(is.na(drinks_content))

```
##Q2
```{r}

drinks_content<-na.omit(drinks_content) %>% 
  select(Beverage_category,Beverage,Beverage_prep,Calories)

drinks_content

```


##Q3
```{r}
drinks_content %>% 
  mutate(calorie_diff = Calories - 135,.after = Beverage) %>% 
  arrange(calorie_diff) %>% 
  select(calorie_diff,Beverage)

```
##Q4
```{r}
max_cal<-drinks_content %>%
  group_by(Beverage_category) %>% 
  summarise(avg_cal = mean(Calories))

max_cal %>% 
  arrange(desc(avg_cal))


min_cal<-drinks_content %>%
  group_by(Beverage_category) %>% 
  summarise(avg_cal = mean(Calories))

min_cal %>% 
  arrange(-desc(avg_cal))

```
##Smoothies, Frappuccino® Blended Coffee and Signature Espresso Drinks contain the highest average calorie counts with 282.2, 276.9, and 250.0, respectively.

### Coffee (4.25), Shaken Iced Beverages (115.9) and Classic espresso drinks (140.2) showed the lowest average calorie counts.

##q5
```{r}
bev_prep_avg<-drinks_content
  
  bev_prep_avg %>% 
    group_by(Beverage_prep) %>% 
    mutate(avg_prep_cal = mean(Calories)) %>% 
    select(Beverage_prep,avg_prep_cal)
  

```
##Q6
```{r}

bev_prep_avg

bev_prep_avg %>%
  mutate(all_drink_avg = mean(Calories)) %>% 
  select(all_drink_avg,Calories,Beverage_prep) %>% 
  filter(Calories > all_drink_avg) %>% 
  group_by(Beverage_prep)



```

##could not get output to combine by groups to neaten display


```{r}
drinks_content %>% 
  filter(Calories == 0)
```

##unsure of combnination Q, surely 0 calories is desired choice
