Hypothesis tests & probability - homework

Question 1.
Explore the dataset and familiarise yourself with it.
```{r}
library(tidyverse)
library(infer)
sleep <- msleep
glimpse(sleep)
head(sleep)

```
Question 2.
Jabberwockies sleep for around 7 hours a night, on average. Perform an appropriate statistical test to determine whether the mean sleep_total in the sampled population of animal types differs from the typical value for jabberwockies.


> does the sleep_total of jabberwockies significantly differ from the 
sample pop (7 hours)?

> H0: no difference in sleep pattern
> H1: the sleep patterns differ

- carrying out two tailed test with P alpha = (<0.05)

```{r}
#calculate observed stat i.e. the mean sleep of the population

sleep_observed_stat <- sleep %>% 
  summarise(mean_sleep = mean(sleep_total))

sleep_observed_stat
```

```{r}
# assess null distribution - testing means which is a point value

sleep_null_dist <- sleep %>% 
  specify(response = sleep_total) %>%
  hypothesise(null = "point", mu = 7) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = 'mean')
```

```{r}


sleep_null_dist %>% 
  visualise(bins = 30) +
  shade_p_value(obs_stat = sleep_observed_stat, direction = "both") # compares
# null mean (7), with observed stat
```
Red line - e.g. observed stat very far from examined value

Assess significance of difference

```{r}
jaber_p_test <-  sleep_null_dist %>% 
  get_p_value(obs_stat = sleep_observed_stat, direction = "both")

jaber_p_test
```

> P value is 0 - we have strong evidence to reject the H0 and state that jaberwockie sleep patterns significantly differ from the sampled animal population.


Question 3.
Perform an appropriate statistical test to determine whether omnivores sleep for significantly longer than herbivores, on average.

> H0: Omnivores and Herbivores do not differ in sleep duration
> H1: Omnivores sleep significantly longer than Herbivores - on average

Q is stating longer - hence one sided, so going for one-tailed 


```{r}
sleep_diet_type <- sleep %>% 
  filter(vore == c("omni", "herbi")) %>% 
  group_by(vore)
```
visualise diference 

```{r}
sleep_diet_type %>% 
  ggplot(aes(y = sleep_total, x = vore)) +
  geom_boxplot()
```
Although herbivore has high variability in sleep pattern, means appears higher so indicative that they sleep longer than omnivores, on average.


```{r}
diet_null_distribution <- sleep_diet_type %>% 
  specify(sleep_total ~ vore) %>% 
  hypothesise(null = "independence") %>% # null is there is no relationship
  generate(reps = 10000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))
```

```{r}
observed_stat_diet <- sleep_diet_type %>% 
  specify(sleep_total ~ vore) %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))

observed_stat_diet
```

Difference in mean sleep pattern between omnivores and herbivores is -0.8957983. Now testing if this is significantly different from 0 difference.

Sticking with right/one tailed as Q asked if omni is greater than herbi sleep - not is there a difference either side of mean.

```{r}
diet_null_distribution %>% 
  visualise() +
  shade_p_value(obs_stat = observed_stat_diet, direction = "right")
```
```{r}
diet_null_distribution %>% 
  get_p_value(obs_stat = observed_stat_diet, direction = "right")
```
> P value > 0.05 therefore we cannot reject the null and can state that on average omnivores do not sleep significantly longer than herbivores.


Question 4. Perform an appropriate statistical test to determine whether the proportion of domesticated animal types in the population of animal types is greater than 5%

Create new col with is domesticated or not and remove NAs
```{r}
sleep <- sleep %>% 
  mutate(is_domesticated = if_else(conservation == "domesticated", "domesticated", "wild"))

sleep_clean <- sleep %>% 
  filter(!is.na(is_domesticated))
```

> H0: the proportion of domesticated animal types is not significantly greater than 5%
> H1: it is signif. greater than 5%

```{r}
domesticated_null_dist <- sleep_clean %>% 
  specify(response = is_domesticated, success = "domesticated") %>% 
  hypothesise(null = "point", p = 0.05) %>% 
  generate(reps = 1000, type = "draw") %>% 
  calculate(stat = "prop")
```

```{r}
domesticated_observed_stat <-  sleep_clean %>%
  specify(response = is_domesticated, success = "domesticated") %>% 
  calculate(stat = "prop")

domesticated_observed_stat
```
0.18 pretty high relative to 5% so indicating signif diff. 

```{r}
domesticated_null_dist %>% 
  visualise(bins = 30) +
  shade_p_value(obs_stat = domesticated_observed_stat, direction = "both")
  
```

```{r}
domesticated_null_dist %>% 
  get_p_value(obs_stat = domesticated_observed_stat, direction = "both")
```
P value is 0, ergo below alpha so have strong evidence that the proportion of domesticated animals in the sampled population is significnatly greater than 5%.


1.2 Hypothesis testing - Defining the Hypothesis

# Q1
Sample population - either do or do not know of shop - could do draw if using string statement or bootstrap if made binary - testing if greater so could use one-tailed ("right").

H0: know of shop != 40%
H1: know of shop >= 40%

H0: the % of people that know the shop is not different from 40%
H1: the % of people that know the shop is different from 40%

# Q2
Two-tailed sample test using null distribution and permute

H0: The CTR of the site with the banner at the top is the same or lower than the CTR of the site with the banner on the right.
H1: CTR on with banner on top is higher than site with banner on righ


H0: CTR at top =< CTR at right
H1: CTR at top > CTR at right

# Q3
Assess mean difference (mu =), one tailed test using bootstrapping.

H0: weekly part width mean same as 145mm
H1: weekly part width mean differs from 145mm

H0: width = 145
H1: width != 145

1.3

# Q1 
Do not reject null, P value greater than alpha so indicative proportion of people
in town having heard of the shop is not higher than 40%.

# Q2
P value below alpha so can reject the null inferring that locating the CTR on sites 
with the banner at the top is greater relative to sites with the CTR on the right.

# Q3
P value higher than alpha so cannot reject the null inferring that the mean width of
does not differ from 14mm.
