---
title: "HW 18_07"
output: html_notebook
---

##Q1 Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.
```{r}
library(tidyverse)
library(dplyr)
library(stringr)

tweet <- read_csv("data/code_clan_tweets.csv")

dim(tweet)

ls(tweet)


```
## 234 rows, 27 columns


##Q2 Find the total number of favourites (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
tweet %>% 
  select(favorite_count,is_quote) %>% 
  filter(is_quote == F) %>% 
  summarise(sum(favorite_count))
```
##425


##Q3 Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.
```{r}

tweet%>% 
  filter(is_quote == F) %>% 
  group_by(source) %>% 
  summarise(mean(retweet_count))

  
```

##Q4 Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

```{r}


tweet_media_adjusted <- tweet %>% 
                  mutate(media_type = replace_na(media_type, "text")) %>% 
                  group_by(media_type) %>% 
                  summarise(sum(favorite_count))
                

tweet_media_adjusted

```

##Q5 Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).

```{r}
tweet %>% 
  summarise(mean(display_text_width))
#149 
#doesnt include twitter handles


tweet %>% 
  mutate(letter_count = str_count(text, ".")) %>% 
  summarise(mean(letter_count))
#188
#doesnt with twitter handles

tweet %>% 
  mutate(letter_count = str_length(text)) %>% 
  summarise(mean(letter_count))
#188



##or 

tweet %>% 
  mutate(letter_count = str_length(text)) %>% 
  summarise(mean(letter_count))

tweet %>% 
  select(display_text_width,text)


tweet %>% 
  summarise(sum(display_text_width))

34964 / 234

ls(tweet)

```
## values differ 148.42 and second line returns 188.37
## 2nd line appears correct, not sure if spaces included in str_count??

#Q6 The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.
```{r}

info <- read_csv("data/code_clan_info.csv")

join_tweet <- tibble(full_join(tweet, info, "tweet_id"))

```

## full join keeps all data



##Q7 %>% From your new joined data, create a new tibble codeclan_hashtags containing only tweet_id and hashtags, then convert hashtags into lowercase for analysis. Keep only those tweets with hashtags.

```{r}

codeclan_hashtags <- join_tweet %>% 
    select(tweet_id, hashtags)

codeclan_hashtags <- tibble(codeclan_hashtags) %>% 
  na.omit(hashtags) %>% 
  mutate(hashtags = str_to_lower(hashtags))
codeclan_hashtags

```

##Ext Q 8
```{r}
library(stringr)

pattern <- "^c\\("

hash <- codeclan_hashtags$hashtags
# could not find hastags variable so made into vector, unsure if this 
# is best way

count <- str_detect(hash, pattern)

str_extract(hash, pattern )

count <-str_count(hash, pattern)
sum(count)

#or

str_sub(hash, 1, 2 )



```

##Q9 Use the str_detect() function to find all tweets with text that mentions “Edinburgh”, and count how many such tweets there are.

```{r}

## used lower case as changed earlier
str_detect(hash, "edinburgh")

str_count(hash, "edinburgh")
ed_count <- str_count(hash, "edinburgh")
sum(ed_count)

hash


```

# 8 edinburgh entries found



#Q10 Use str_extract_all() with a regular expression to find out which Twitter users CodeClan have been tweeting.
```{r}

tweet_text <- join_tweet %>% 
  select(text)

handles <- tweet_text$text

handle_hits <-str_extract_all(handles, "(?<!\\w)@[A-Za-z0-9_]+") %>% 
  flatten_chr()

unique(handle_hits)

handle_sum <- tibble(handle_hits)

handle_sum %>% 
    count(handle_hits) %>% 
    arrange(desc(n))
  
```

