---
title: "Manual model development"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', warning = FALSE, message = FALSE)
```


# Tidy up the data ready for regression:
```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
```

```{r}
house <- read_csv("kc_house_data.csv")
head(house)
summary(house)
```
# Removal unnecessary variables
```{r}
house <- house %>% 
  select(-c(id, date, sqft_living15, sqft_lot15, zipcode))
```

# Convert variables where needed
```{r}
str(house)

house <- house %>% 
  mutate(waterfront = as.logical(waterfront),
         view = as.factor(view),
         condition = as.factor(condition),
         grade = as.factor(grade),
         renovated = ifelse(yr_renovated > 0, T, F)) %>% 
  select(-yr_renovated)
  
```
# Check for aliased variables 
```{r}
alias(price ~ ., data = house)
```

Sqft_living id'd as an alias for sqft_above and therefore dropped from model
```{r}
house <- house %>% 
  select(-sqft_living)
  
```

Systematically build a regression model containing up to four main effects 
```{r}
# look at numeric variables
house_numeric <- house %>% 
  select_if(is.numeric)

ggpairs(house_numeric)
```

Price will be my model's response variable, bathroom has good coeff (0.5) and shows
evidence of regression trend in dotplot, also sqft_above (0.6) and sign of relationship.

Planning on log transforming price as vary massively 
```{r}
house %>% 
  summarise(
    min = min(price),
    max = max(price)
  )
```



Now looking at categorical variables
```{r}
house_categorical <- house %>% 
  select_if(function(x) !is.numeric(x)) %>% 
  bind_cols(house$price)

```

```{r}
ggpairs(house_categorical)
```
Waterfront looks well staggered between price, grade also shows relationship


Build some plots with these variables 
```{r}
house %>% 
  ggplot(aes(x = log(price), y = sqft_above)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```
Pretty clear linear trend

Model attempt v.0.1
```{r}
model1 <- lm(log(price) ~ sqft_above, data = house)
summary(model1)

```
```{r}
autoplot(model1)
```
Significant relationship between sqft_above and price, diagnostics look close to 
ideal.

R2 is 36% - continue to try and bump up

Next; adding bathrooms

```{r}
ggplot(house, aes(x = as.factor(bathrooms), y = price)) +
  geom_boxplot()

# 8 bathrooms? aggressive. 
```

```{r}
model2 <- lm(log(price) ~ sqft_above + bathrooms, data = house)
summary(model2)
```
```{r}
autoplot(model2)
```
Same story as sqft_above, only adds 3% onto R2 so not much addition. Will continue 


Attempt 3 - Waterfront
```{r}
model3 <- lm(log(price) ~ bathrooms + waterfront + sqft_above, data = house)

summary(model3)
```

```{r}
autoplot(model3)
```

All significant and good diagnostics, adding grade as seems logical predictor of
house price and not too high coeff

```{r}
model4 <- model3 <- lm(log(price) ~ bathrooms + waterfront + sqft_above + grade, data = house)
```

```{r}
summary(model4)
```
```{r}
autoplot(model4)
```

Higher graded house have significant effect on house prices, model is at 53% pretty good.
Diagnostics look good on the whole so sticking with final model 

lm(formula = log(price) ~ bathrooms + waterfront + sqft_above + 
    grade, data = house)