```{r}
library(dplyr)
library(tidyverse)
library(skimr)
movies<-read_csv("movies.csv")
movies
skim(movies)
movies<-movies
unique(movies)
names(movies)

```
##Q1
- remove NAs, impute/remove
- fix column names, clean with janitor



##Q2 

```{r}
 
##subset desired coolumns
movies_selected<-movies %>% 
  select(title,runtime,budget)

##identify NAs and whihc col
movies_selected %>% 
  summarise(across(everything(),~sum(is.na(.))))

##or anywhere
movies_selected %>% 
  summarise(count = sum(is.na(.)))

```
## 2 missing values in runtime, 0 in title and budget.

##Q4
```{r}
##change 0 runtime values to NAs
movies_runtime<-movies_selected %>%
  mutate(runtime = na_if(runtime, 0))

##check NAs are found
movies_runtime %>%
  filter(is.na(runtime))

```

##Q5 change na values to median value with coalesce and remove NA
```{r}
movies_imputed<-movies_runtime %>%
  mutate(runtime = coalesce(runtime, median(runtime,na.rm = T)))
  
movies_imputed %>%
  filter(is.na(runtime))

```
##Q6 find 10 shortest/longest movies
```{r}
movies_imputed %>% 
  slice_min(runtime, n = 10)

movies_imputed %>% 
  slice_max(runtime, n = 10)
```
##Q7 Using the if_else function, impute anything with a budget below $100 with the median budget. Overwrite your movies_imputed so it contains the new budget values.


```{r}

median_budget<-median(movies_imputed$budget)
median_budget

movies_imputed<-movies_imputed %>% 
  mutate(budget = if_else(budget < 100, median_budget,budget))

##or - but this took longer to figure out - 

movies_imputed<-movies_imputed %>% 
  mutate(budget = if_else(budget < 100, median(budget),budget))
         


```

##QE1
Using case_when, create a new column called budget_type and label budgets as follows:

Budgets below 12 million as ‘Small budget’
Budgets between 12 and 40 million as ‘Medium budget’
Budgets 40 millions and above as ‘Big budget’
Save this into a new object called movie_budgets.
```{r}
         
   movie_budgets <-movies_imputed %>% mutate(budget_type = 
                         case_when(budget >= 100 & budget < 12e6   ~ "Small Budget", 
                                   budget > 12e6 & budget < 40e6 ~ "Medium Budget",
                                   budget >= 40e6 & budget <= 380e6 ~ "Big Budget", TRUE ~ "Oth"))
     
                                  
   
```   

##usure with this, struggled doing ranges of values and what %% == 0 etc is in syntax


##QE2 -- brain no work now

