---
title: "Simple linear regression homework"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', warning = FALSE, message = FALSE)
```

```{r}

library(tidyverse)
library(GGally)

project <- read_csv("project_management.csv")

```
# Plot the data
```{r}

project %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(
    y = "Actual Length",
    x = "Estimated Length"
    
  )
```

# Calculate the correlation coefficient 
```{r}

project %>% 
  ggcorr(label = T)

```
Correlation coefficient between actual length and estimated length is 0.8;
indicating a very strong positive correlation between these variables. 


# Perform a simple linear regression

```{r}
model1 <- lm(actual_length ~ estimated_length, data = project)

model1

summary(model1)
```
Estimated length has a significant relationship with actual length allowing
actual length to be calculated using estimated length. The coeff of 1.2235 tells us
for every increase in day by 1 increases the actual length by 1.2235 = Jobs are 
taking longer than expected. The R2 is high (64%) so this the variation in the 
actual length is well explained by the estimated length.

# Is the relationship statistically significant?

```{r}
# check diagnostics
library(ggfortify)

autoplot(model1)
```

# Diagnostics

Residuals look independent and well spread above/below 0 intercept, QQ plot looks 
normally distributesalong the slope, with the exception of a couple out outlier 
shouldn't be an issue as sofew. Scale location plot show no funneling so infers 
homoscedasticity. Collectively this indicates good model fit of the data. 

```{r}
cor.test(project$estimated_length, project$actual_length, method = c("pearson"))
```


Pearson's correlation confirms relationship is statistically significant (p < 0.05).