---
title: "HW_22_08"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

1.
```{r}
library(tidyverse)
library(CodeClanData)

backpack

head(backpack)
names(backpack)
```

2. What is the distribution of backpack weights that students are carrying? 
```{r}
backpack %>% 
  ggplot(aes(x = backpack_weight)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(x = "Backpack weights ",
       y = "Frequency")

shapiro.test(backpack$backpack_weight)

backpack %>% 
  summarise(mean(backpack_weight))


```
Hist plot shows majority of backpack weights arounf 5 - 12 (kg?). One kid has super heavy bag around 35kg. The distribution of backpack weights is non-normal. Average bag weight is 11.66kg

3. Are there differences in backpack weight between sex?
```{r}
backpack %>% 
  ggplot(aes(x = sex, y = backpack_weight, fill = sex)) +
  geom_boxplot() +
  labs(title = "Backpack weight between genders",
       x = "",
       y = "Backpack weight")
```
Average backpack weight between genders is very similar, male bags are just marginally heavier on average. Males also have a greater spread in backpack weights largely due to the one extreme outlier with a 35kg bag. 

4. What is the relationship between body weight and backpack weight?
```{r}

backpack %>% 
  ggplot(aes(x = body_weight, y = backpack_weight)) +
  geom_point() +
  labs(title = "Backpack weight between genders",
       x = "Body weight",
       y = "Backpack weight") +
  geom_smooth(method = "lm")



shapiro.test(backpack$body_weight)
shapiro.test(backpack$backpack_weight)
#data non normal so using spearman 

 
  cor.test(backpack$body_weight, backpack$backpack_weight, method = "spearman")

```

Very slight correlation between body weight and backpack weight (Cor.coeff = 0.21 (p < 0.05)). Minor trend of greater body weights have heavier backpacks. 

5.
```{r}
backpack %>% 
  ggplot(aes(x = body_weight, y = backpack_weight, color = sex)) +
  geom_point() +
  labs(title = "Backpack weight between genders",
       x = "Body weight",
       y = "Backpack weight") +
  facet_wrap(~ sex)


```

Females have lower body weights but have bags similiar in weight to males.Individual with heaviest bag is male, overall scatter does not indicate influence of gender. 

6.What proportion of students report back pain?
```{r}
names(backpack)
backpack_pain <- backpack %>% 
  mutate(back_status = ifelse(back_problems == 1, "yes", "no"))

backpack_pain %>% 
ggplot(aes(x = 1, fill = back_status)) +
  geom_bar(position = "stack") +
  labs(x = NULL,
       y = "% Pain",
       fill = "Back pain reported") +
    theme(legend.position = "right",
        axis.text.x = element_blank(),
        axis.title.x = element_blank()) 

```
Majority of students don't report back pains, over a quarter do.

7.
```{r}
  names(backpack)

backpack_pain %>%
  ggplot(aes(x = sex, fill = back_status, label = stat(count))) +
  geom_bar(position = "stack") +
  labs(x = "Sex",
       y = "Count",
       fill = "Pain Status") +
  theme_minimal() +
  theme(legend.position = "right")
```
More girls than boys in study pool. Females had more reports of bakpain than males by almost double. 

8. Females show higher prevalence of backpain than males. 
```{r}

ggplot(backpack_pain) + 
  aes(x = body_weight, y = backpack_weight, color = back_status) + 
  geom_point() + 
  facet_grid(sex ~ back_status) + 
  labs(x = "Body Weight", 
       y = "Backpack Weight",
       color = "Reported back pain")+
  theme(legend.position = "right") +
  geom_smooth(method = "lm", se = F)


```

Ex 1.
```{r}
b_count <- backpack_pain %>% 
  group_by(year, back_status) %>% 
  summarise(count = n()) %>% 
    filter(year != "0")
  ungroup()


b_count %>% 
  ggplot(aes(x = year, y = count, fill = back_status)) +
  geom_bar(stat = "identity",position = "dodge", width = 0.6) +
  labs(
    x = "Year",
    y = "Back pain report",
    fill = "Reported back pain"
  ) +
  theme(legend.position ="right")
```
Back pain reports peak in year 3 and decrease thereafter, reports are lowest in early years. 

Ex. 2
```{r}


sample_sizes <- backpack %>%
  group_by(year) %>%
  summarise(sample_size = n()) %>%
  ungroup()

backpack %>% 
  filter(year != 0) %>% 
  ggplot(aes(x = backpack_weight)) +
  geom_bar() +
  labs(
    x = "Year"
  ) +
  facet_wrap(~ year)
```

More females than males reported back pain, females show slight positive correlation between body weight and back pack weight. This trend was also seen in those who didn't report back pain. In contrast, males who did report back pain showed a decreasing backpack weight with increasing body weight however the sample size in this group is minimal (6) so no real conclusions can be made. 