---
title: "Homework - features and elements of multiple regression"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', warning = FALSE, message = FALSE)
```


1.
```{r}
library(tidyverse)

prices <- read_csv("housing_prices.csv")


head(prices)
```

2. 

0.9 correlation coeff is super high showing variables are strongly correlated, bottom left plot shows this also
```{r}
library(GGally)

prices %>% 
  select(total_bedrooms, total_rooms) %>% 
  ggpairs()

```

3.
```{r warning=FALSE}
prices <- prices %>% 
  select(-total_bedrooms)
```

4.
```{r warning=FALSE}
ggpairs(prices , progress = F,
        lower = list(combo = wrap("facethist", bins = 15)))
```
strong correlation: median house value ~ median income
correlation: median house value ~ latitude
ocean prox has staggering

```{r}
prices %>% 
  ggplot(aes(median_income, median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
prices %>% 
  ggplot(aes(latitude, median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
prices %>% 
  ggplot(aes(ocean_proximity, median_house_value)) +
  geom_boxplot()
```

5.(n - 1): 5 groups so 4 dummies

6. Quite a lot of residuals above 0 intercept. Tails on qq not good, funneling heteroscedasticity; 
overall not a great model, moderately skewed
```{r}
library(ggfortify)

model1 <- lm(median_house_value ~ median_income, data = prices)

model1

autoplot(model1)
```

```{r}
summary(model1)
```

```{r}

model2 <- lm(median_house_value ~ median_income + ocean_proximity, data = prices)

model2

autoplot(model2)
```

```{r}
summary(model2)
```
Model explains 55% of variation in median house value, island locations have
highest influence on prince, inland reduces value. 

8.
```{r}
model3 <- lm(log(median_house_value) ~ log(median_income) + ocean_proximity, data = prices)

model3

autoplot(model3)

summary(model3)
```

Log transformation improves model R2. 


```{r}

prices %>% 
  ggplot(aes( x = log(median_income), y = log(median_house_value))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ ocean_proximity)

```

